Chemical tagging can work: \\ Identification of stellar phase-space structures \\ purely by chemical-abundance similarity

### 2016-01-18 papers/clusters.tex
notes to self
notes to self, figure

% to-do
% - write discussion section w MKN input
% - expand kmeans discussion
% - put in references
% - cite codes!
% - fix all remaining occurrences of ARC, MKN, HWR, DWH, or CITE
% - forward to Freeman, others?
% - submit to journal and arXiv

@@

% results shit
\newcommand{\totalnumber}{98462} %% DWH CHECK

@@

% ARC asks DWH: Why aren't we showing The Cannon Teff/logg in these
% 				figures? Even if they aren't used for clustering..

@@

Another issue with k-means is that it effectively uses metric
distances in the $D$-space; it presumes Euclidean isotropy.
We choose here to work in the hydrogen-normalized abundance space, the
space of [C/H], [N/H], [O/H], [Na/H], [Mg/H], [Al/H], [Si/H], [S/H],
[K/H], [Ca/H], [Ti/H], [V/H], [Mn/H], [Fe/H], [Ni/H].
But in addition to this, we re-scale these by approximate measurement
precisions obtained by \thecannon\ before running k-means.
This makes the space close to isotropic in measurement uncertainty or
observational precision.
Importantly we use \emph{only} abundance-space information, and no
positional or velocity information (nor $\teff$ nor $\logg$ nor any
targeting or observational meta-data) as input to the clustering
algorithm.

@@

When the k-means results are returned, the $D\times D$ empirical
variance tensor for the members of each cluster can be constructed,
and an effective density can be computed as the number of points in
the cluster divided by the square-root of the determinant of the
tensor.
This density was used to rank abundance-space overdensities (from highest
to lowest) for visual inspection.
We chose a few interesting cases from many and show them in
\figurename~\ref{fig:M13}, \ref{fig:Pal5}, \ref{fig:Sgr},
\ref{fig:halo}, and \ref{fig:disk}.
The first three of these are known clusters M13, Pal5, and Sagittarius (MKN, ARC, help!).
The fourth is a halo structure with high velocity dispersion and possibly accreted.
The fifth is a thin-disk star-formation feature.
We will discuss these more below.

DWH \& HWR: How do we show that we aren't just getting clusters because \apogee\ has plate-to-plate calibration variations?

\begin{figure}[!p]
\insanefigure{cluster_0256_0253}
\caption{Same as \figurename~\ref{fig:all} except that the entire
  sample has been made gray and the members of the most dense
  $K=256$ abundance-space clusters have been rendered as unique, prominent triangles.
  This cluster, which was identified only in abundance space (six
  projections of which are the top six panels in this \figurename),
  turns out to be dominated by the halo globular cluster
  M13.\label{fig:M13}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0034}
\caption{Same as \figurename~\ref{fig:M13} but for another dense
  abundance-space cluster.
  This one turns out to be dominated by the disrupting globular
  cluster Palomar 5.\label{fig:Pal5}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0177}
\caption{Same as \figurename~\ref{fig:M13} but for another dense
  abundance-space cluster.
  This one turns out to be dominated by the Sagittarius star
  cloud.\label{fig:Sgr}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0010}
\caption{Same as \figurename~\ref{fig:M13} but for another dense
  abundance-space cluster.
  This one turns out to be dominated by a high-velocity-dispersion
  structure in the Galaxy halo.\label{fig:halo}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0141}
\caption{Same as \figurename~\ref{fig:M13} but for another dense
  abundance-space cluster.
  This one turns out to be dominated by a very thin young stellar
  structure in the Galaxy disk.\label{fig:disk}}
\end{figure}

>>>>

% to-do
% - write discussion section w MKN input
% - fix all remaining occurrences of ARC, MKN, HWR, DWH, or CITE
% - submit to journal and arXiv
% - forward to Freeman, Bovy, others?

@@

% results shit
\newcommand{\totalnumber}{98462}

@@

ARC asks DWH: Why aren't we showing The Cannon Teff/logg in these
				figures? Even if they aren't used for clustering..

@@

Another issue with k-means is that it effectively uses metric
distances in the $D$-space; it presumes Euclidean isotropy.
We choose here to work in the hydrogen-normalized abundance space, the
space of [C/H], [N/H], [O/H], [Na/H], [Mg/H], [Al/H], [Si/H], [S/H],
[K/H], [Ca/H], [Ti/H], [V/H], [Mn/H], [Fe/H], [Ni/H].
But in addition to this, we re-scale these by approximate typical measurement
precisions obtained by \thecannon\ before running k-means.
This makes the space close to isotropic in measurement uncertainty or
observational precision.
Importantly we use \emph{only} abundance-space information, and no
positional or velocity information (nor $\teff$ nor $\logg$ nor any
targeting or observational meta-data) as input to the clustering
algorithm.

@@

When the k-means results are returned, the $D\times D$ empirical
variance tensor for the members of each cluster can be constructed,
and an effective density can be computed as the number of points in
the cluster divided by the square-root of the determinant of the
tensor.
This density was used to rank abundance-space overdensities (from highest
to lowest) for visual inspection.
In \figurename~\ref{fig:densities} we show the distribution of cluster membership
and density for the $K=256$ run of k-means.

We chose a few interesting cases from the high-density clusters in the $K=256$ run and show them in
\figurename~\ref{fig:M13}, \ref{fig:Pal5}, \ref{fig:Sgr},
\ref{fig:halo}, and \ref{fig:disk}.
The first three of these are dominated by stars in the known clusters M13, Pal5, and Sagittarius.
The fourth is a halo structure with high velocitay dispersion and possibly accreted.
The fifth is a thin-disk star-formation feature.
We will discuss these more below.

DWH: The clusters shown in \figurename~\ref{fig:M13}, \ref{fig:Pal5},
and \ref{fig:Sgr} might not look dense in abundance space, but they
are very dense in the 15-dimensional space.  Indeed,
\figurename~\ref{fig:M13} is the densest cluster found in the 15-d
space \emph{by far} (see \figurename~\ref{fig:densities}).

DWH: The figures show results only from the $K=256$ run, why?

DWH: Even if $K$ was chosen very carefully, it would still break up
real clusters into many clusters, why?  Abundance spreads!

DWH: The algorithm sees many different clusters; the figures show only
a few.  Do we see all clusters we \emph{should} see?  That question is
beyond the scope of this \documentname.

DWH: Each cluster figure and the disk figure shows stars that are part
of the structure and some that are not.  Every cluster will have
background contamination, why?

\begin{figure}[!p]
\includegraphics[width=2\figurewidth]{./figs/clusters_0256.pdf}
\caption{The distribution of membership and density for the 256
  abundance-space clusters returned by the k-means algorithm at
  $K=256$.  The densest cluster---which is shown in
  \figurename~\ref{fig:M13}---is far more dense than the second
  densest.\label{fig:densities}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0253}
\caption{Same as \figurename~\ref{fig:all} except that the entire
  sample has been made gray and the members of the most dense
  $K=256$ abundance-space clusters have been rendered as unique, prominent triangles.
  This cluster, which was identified only in abundance space (six
  projections of which are the top six panels in this \figurename),
  turns out to be dominated by the halo globular cluster
  M13.\label{fig:M13}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0034}
\caption{Same as \figurename~\ref{fig:M13} but for another dense
  abundance-space cluster.
  This one turns out to be dominated by the disrupting globular
  cluster Palomar 5.\label{fig:Pal5}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0177}
\caption{Same as \figurename~\ref{fig:M13} but for another dense
  abundance-space cluster.
  This one turns out to be dominated by the Sagittarius star
  cloud.\label{fig:Sgr}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0010}
\caption{Same as \figurename~\ref{fig:M13} but for another dense
  abundance-space cluster.
  This one turns out to be dominated by a high-velocity-dispersion
  structure in the Galaxy halo.\label{fig:halo}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0141}
\caption{Same as \figurename~\ref{fig:M13} but for another dense
  abundance-space cluster.
  This one turns out to be dominated by a very thin young stellar
  structure in the Galaxy disk.\label{fig:disk}}
\end{figure}
